---
title: "Ch2. Standard Statistical Inference"
author: "Jenny_Lee"
date: "January 29, 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries to be used in our analysis.
```{r libraries}
library(MASS)
library(ggplot2)
library(stringr)
```



## Dataset for exercises


```{r dataset}

set.seed(123)
x1 <- trunc(rlnorm(n = 753, 5)) # log normal distribution with logmean = 5
x2 <- rgamma(n = 267, 1.2, .25)
# fit <- fitdistr(x2, "lognormal")

```

## 2.1 Probability functions for Benford's law

### 1) Probability density function 

```{r dBenford}

dBenford <- function(n) {
  # n: any number or a list of numbers 
  if (length(n) == 1){
    # If a number is provided,
    if (n %in% 1:9) {
    # If  n is element of the sample space
      log(1 + 1 / n) / log(10)
    } else {
      0
  }} else{
    # if more than one numer is provided, outputs the corresponding list 
    tmp_list <- lapply(n, dBenford)
    unlist(tmp_list)
  }
}

# Densities of {1, 2, ..., 8, 9}
dBenford(1:9)

# The sum of density on the sample space is equal to 1.
sum(dBenford(1:9))
```


### 2) Cumulative distribution function 

```{r cBenford}

pBenford <- function(n){
  # n: a numer or a list of numbers 
  if (length(n) == 1){
    input_space <- 1:9
    selection <- input_space[input_space <= n]
    sum(dBenford(selection))
  } else {
    # If more than one number is provided,
    tmp_list <- lapply(n, pBenford)
    unlist(tmp_list)
  }
}

pBenford(1:9)


```

### 3) Random number generator 

```{r random}

rBenford <- function(n = 100){
  # n: integer that is greater than 1 
  cdf <- pBenford(1:9)
  if (n == 1){
    r <- runif(1)
    length(cdf[cdf < r] + 1) 
  } else {
    tmp_list <- lapply(1:n, function(x){
      r <- runif(1)
      length(cdf[cdf < r]) + 1 })
    unlist(tmp_list)

  }
}

```


Let's check if our *rBenford* function generate random numbers according to the Benford's law. 

```{r sim}

sim_res <- lapply(1:100, function(x){rBenford(n = 100)})
sim_res <- data.frame(res = unlist(sim_res))
dBenfords <- dBenford(1:9)
small_data <- data.frame(n = 1:9, density = dBenfords)
ggplot(data = sim_res, aes(x = res, y = ..density..)) + 
  geom_histogram(bins = 9, colour = "grey60", fill = "white") + 
  geom_line(data = small_data, aes(x = n, y = density)) + 
  geom_point(data = small_data, aes(x = n, y = density), 
             shape = 22,  size = 4, colour = "darkred", fill = "pink") + 
  scale_x_discrete(limits = 1:9) + 
  labs(title = "Histogram of simulation (n = 10,000) Vs. Real density of Benford's Law") + 
  xlab("sample space")

```




## 2.2  

#### Given a vector of integers, x, write a function that counts frequency of first digits of elements of x. Write a function, based on Pearson's statistics, that tests if Benfords distribution is relevant. Try this function on X1. 

```{r first_digit_freq}

first_digit <- function(n){
  # First convert the provided list of numbers to string
  n <- strsplit(toString(n), split = ", ")
  first_digits_list <- lapply(1:length(n[[1]]), function(x){
    # Take each number then retrieve the first digit only
    strsplit(n[[1]][x], split = "")[[1]][1]
  })
  # Counting frequency
  table_tmp <- table(as.numeric(unlist(first_digits_list)))
  df = data.frame(table_tmp)
  colnames(df) <- c("first_digit", "freq")
  df
}

first_digit(1:1000) 
```





